\sysfull

### 2017-08-01 paper/background.tex
makefile patch that should build figs and background edits

Software has different behavior when executed intermittently than when executed
with continuous power.  Each power failure clears volatile memory (registers,
stack, and global variables). Non-volatile memory (e.g., FRAM) persists across
power failures. On a power failure, control flows back to a prior point in the
execution. By default, after a power failure control flows to the beginning of
{\tt main()}. Early work on intermittent computing preserve intermittent
progress by periodically capturing and restoring to a non-volatile checkpoint
of the volatile execution context (i.e., registers and
stack)~\cite{mementos,hibernusplusplus,quickrecall,idetic}

@@

%\TODO{This sequential/modular split seems pretty odd to me. I think we can just talk about tasks in the sense of ratchet, alpaca, and chain.  Anything with static (programmer-written or compiler inserted) boundaries applies}

Intermittent execution models combatting WAR can be classified into: \emph{sequential execution model} (which historically formed a baseline for this stream of research) and \emph{modular execution model} (a potent alternative, on which we base the design decisions of \sys). \todo{Remove division per sequential/modular split}{Przemek}

Under this model a program for intermittently-powered device is seen as a one idempotent operational region with one common context. Generally, the progress of the program is saved and updated by means of checkpointing, refer again to Section~\ref{sec:background_consistency}. Normally, the sequential model relies on a hardware assistant to measure the voltage level in the energy reservoir to place a checkpoint~\cite{mementos,mottola2017harvos,hibernus}. The benefit of this model is that it does not require code modification by a programmer. However it suffers from significant overhead (refer to experiment result in~\cite[Fig. 3]{chain}, where it is concluded that each checkpoint causes overhead). \todo{I use the word `checkpoint' for normal checkpoint and task boundary interchangeably---check if it does not cause confusion}{Brandon}

At the heart of this model is the concept of an \emph{idempotent task}. The idempotent task is a C language function that does not have arguments and does not return a value. This task uses a well defined interface to interact with the non-volatile memory. Therefore, it tolerates arbitrary number of power interrupts. This model, generally, produces less overhead~\cite{chain} and allows multiple applications to run on an intermittently-powered device by interleaving their tasks. However, it obviously requires code modification. For example, if an algorithm is written according to the continuous execution model it has to be split by a programmer into small tasks to run under the modular intermittent execution model.

>>>>

Software has different behavior when executed intermittently than when executed
with continuous power.  Each power failure clears volatile memory (registers,
stack, and global variables). Non-volatile memory (e.g., FRAM) persists across
power failures. On a power failure, control flows back to a prior point in the
execution. By default, after a power failure control flows to the beginning of
{\tt main()}. Early work on intermittent computing preserve intermittent
progress by periodically capturing and restoring to a non-volatile checkpoint
of the volatile execution context (i.e., registers and
stack)~\cite{mementos,quickrecall}, sometimes using
hardware support~\cite{mementos,mottola2017harvos,hibernusplusplus,hibernus,idetic}.

@@

%\TODO{This sequential/modular split seems pretty odd to me. I think we can just talk about tasks in the sense of ratchet, alpaca, and chain.  Anything with static (programmer-written or compiler inserted) boundaries applies}
Task-based intermittent execution models~\cite{dino,chain,alpaca} ask the
programmer to decompose their program into a collection tasks. A task is a
function with no caller that contains arbitrary computation, sensing, and
communication.  A programmer describes the sequence in which tasks may execute,
creating a task graph. The flow of control from task to task happens at
programmer demarcated points and may conditionally depend on program values.
Task-based programming abstractions guarantee that tasks exhibit behavior as
though they executed {\em atomically}, regardless of arbitrary power failures.
Task-based runtime systems tolerate failures by repeatedly attempting to
execute a task until it completes. Task-based runtime systems ensure
task-atomic semantics by ensuring that repeated task executions are idempotent
(i.e., have the same effect when repeatedly executed).  The key to idempotence
is ensuring that updates to non-volatile state made by an interrupted, partial
execution of a task are never visible to a subsequent (potentially completing)
re-execution of the task.

There are several run time strategies to ensuring task idempotence. One
approach~\cite{alpaca} is to identify non-volatile data involved in WAR
dependences in a task (like DINO and Ratchet did for
checkpoints~\cite{dino,ratchet}), execute the task using buffered copies of
those data, and commit the buffered copies before completing the task. Another
strategy is to statically create multiple versions of non-volatile data shared
by tasks and ensure by construction that no task reads and writes the same
non-volatile memory location~\cite{chain}.  Regardless of the idempotence
strategy, the key to task-based execution is that fixed, statically defined
tasks execute as though atomic, completing after zero or more attempts.

\TODO{Brandon left off here.}
